Technical Report

### 2017-09-27 fully-abstract-comp/technical-report.tex
program definition (work in progress)

% Math packages
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{array}
\usepackage{cleveref}
\usepackage{stmaryrd}

@@

@@

%%% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\rn}{\var{rn}}
\newcommand{\reg}{\var{reg}}
\newcommand{\mem}{\var{mem}}
\newcommand{\ms}{\var{ms}}
\newcommand{\pc}{\var{pc}}
\newcommand{\stk}{\var{stk}}
\newcommand{\stkf}{\stk_{\var{frame}}}
\newcommand{\ret}{\var{ret}}
\newcommand{\data}{\var{data}}
\newcommand{\code}{\var{code}}
\newcommand{\priv}{\var{priv}}
\newcommand{\opc}{\var{opc}}
\newcommand{\odata}{\var{odata}}
\newcommand{\vsc}{\var{sc}}
\newcommand{\cb}{\var{cb}}
\newcommand{\baddr}{\var{b}}
\newcommand{\eaddr}{\var{e}}
\newcommand{\aaddr}{\var{a}}
\newcommand{\stdrng}{[\baddr,\eaddr]}

@@

%%% Permissions
\newcommand{\plainperm}[1]{\textsc{#1}}
%\newcommand{\rwlx}{\plainperm{rwlx}}
\newcommand{\rwx}{\plainperm{rwx}}
\newcommand{\rx}{\plainperm{rx}}
%\newcommand{\rwlxo}{\plainperm{rwlxo}}
%\newcommand{\rwlo}{\plainperm{rwlo}}
%\newcommand{\rwl}{\plainperm{rwl}}
%\newcommand{\rwxo}{\plainperm{rwxo}}
%\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
%\newcommand{\rxo}{\plainperm{rxo}}
%\newcommand{\ro}{\plainperm{ro}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\noperm}{\plainperm{0}}
%\newcommand{\nopermo}{\plainperm{0o}}
\newcommand{\enter}{\plainperm{e}}
%\newcommand{\entero}{\plainperm{eo}}

@@

\begin{document}
\section{The two capability machines}
\subsection{Domains}
\label{sec:domains}

@@

@@

The source machine is also a capability machine with memory capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine. I could not recall why we had this, so I removed it. (I think it may have been from when we had call and return instructions?)}
\dominique{26-9-2017: I think we should allow $\rstk$ in the source language too.}
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

\subsection{Operational Semantics}
\subsubsection{Notes}
Generally:
\begin{itemize}
%\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
\item Linear capabilities are cleared when they move around in memory.
\end{itemize}

@@

\noindent\textbf{restrict}\\
This instruction uses the $\decPerm{}$ function.
\begin{align*}
  \sem{\trestrict{r_1}{r_2}{r_3}} = &
                                      \begin{cases}
                                        \updPcAddr{}\left(
                                          \arraycolsep=0pt
                                          \begin{array}{rl}
                                          \Phi&\updReg{r_2}{w_2}\\
                                              &\updReg{r_1}{c}
                                          \end{array} \right)
&
                                        \arraycolsep=0pt
                                        \begin{array}[t]{l}
                                          \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                          \quad\text{and in either case $n \in \ints$ and} \\
                                          \decPerm{n} \sqsubseteq \perm \tand \\
                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\
                                          w_2 = \linCons{\Phi(r_2)}
                                        \end{array}
                                        \\
                                        \sourcecolor\left.
                                        \updPcAddr{}\left(
                                          \arraycolsep=0pt
                                          \begin{array}{rl}
                                          \Phi&\updReg{r_2}{0}\\
                                                    &\updReg{r_1}{c}
                                          \end{array}\right)\right.
                                        &
                                        \sourcecolor\left.
                                        \arraycolsep=0pt
                                        \begin{array}[t]{l}
                                          \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                          \quad\text{and in either case $n \in \ints$ and} \\
                                          \decPerm{n} \sqsubseteq \perm \tand \\
                                          c = \stkptr{\decPerm{n},\baddr,\eaddr,\aaddr}
                                        \end{array}\right.
                                        \\
                                        \failed & \totherwise
                                      \end{cases}
\end{align*}

@@

\noindent\textbf{cseal}\\
\begin{align*}
  \sem{\tcseal{r_1}{r_2}{r_3}}(\Phi) = &
                                  \begin{cases}
                                    \arraycolsep=0pt
                                    \begin{array}{rl}
                                      \Phi&\updReg{r_1}{\vsc}\\
                                           \updReg{r_2}{w_2}
                                    \end{array}
&
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) \in \SealableCaps \tand \\
                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
                                      \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
                                      \vsc = \sealed{\Phi(r_2),\sigma} \tand\\
                                      w_2 = \linCons{\Phi(r_2)}                                     
                                    \end{array}
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

@@

We would like splice and split to have the following properties

@@

\subsection{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]

>>>>

% Math packages
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{array}
\usepackage{cleveref}
\usepackage{stmaryrd}

@@

%%% Math environments
\declaretheorem[numbered=yes,name=Lemma,qed=$\blacksquare$]{lemma}
\declaretheorem[numbered=yes,name=Theorem,qed=$\blacksquare$]{theorem}
\declaretheorem[numbered=yes,name=Definition,qed=$\blacksquare$]{definition}
\declaretheorem[numbered=yes,name=Specification,qed=$\blacksquare$]{specification}

@@

%%% Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\rn}{\var{rn}}
\newcommand{\reg}{\var{reg}}
\newcommand{\mem}{\var{mem}}
\newcommand{\ms}{\var{ms}}
\newcommand{\pc}{\var{pc}}
\newcommand{\stk}{\var{stk}}
\newcommand{\link}{\var{link}}
\newcommand{\stkf}{\stk_{\var{frame}}}
\newcommand{\ret}{\var{ret}}
\newcommand{\data}{\var{data}}
\newcommand{\code}{\var{code}}
\newcommand{\priv}{\var{priv}}
\newcommand{\opc}{\var{opc}}
\newcommand{\odata}{\var{odata}}
\newcommand{\vsc}{\var{sc}}
\newcommand{\cb}{\var{cb}}
\newcommand{\baddr}{\var{b}}
\newcommand{\eaddr}{\var{e}}
\newcommand{\aaddr}{\var{a}}
\newcommand{\stdrng}{[\baddr,\eaddr]}

@@

%%% Permissions
\newcommand{\plainperm}[1]{\textsc{#1}}
%\newcommand{\rwlx}{\plainperm{rwlx}}
\newcommand{\rwx}{\plainperm{rwx}}
\newcommand{\rx}{\plainperm{rx}}
%\newcommand{\rwlxo}{\plainperm{rwlxo}}
%\newcommand{\rwlo}{\plainperm{rwlo}}
%\newcommand{\rwl}{\plainperm{rwl}}
%\newcommand{\rwxo}{\plainperm{rwxo}}
%\newcommand{\rwo}{\plainperm{rwo}}
\newcommand{\rw}{\plainperm{rw}}
%\newcommand{\rxo}{\plainperm{rxo}}
\newcommand{\readonly}{\plainperm{r}}
\newcommand{\ro}{\readonly}
\newcommand{\noperm}{\plainperm{0}}
%\newcommand{\nopermo}{\plainperm{0o}}
\newcommand{\enter}{\plainperm{e}}
%\newcommand{\entero}{\plainperm{eo}}

@@

\newcommand{\range}[1]{\plainfun{range}{#1}}

\title{Technical Report}

\begin{document}
\maketitle
\tableofcontents
\section{The two capability machines}
\subsection{Domains}
\label{sec:domains}

@@

\subsubsection{Useful definitions}
\begin{definition}
  For a capability $c=((\_,\_),\baddr,\eaddr,\_)$ we say it has range $[\baddr,\eaddr]$ and we define
  \[
    \range{c} = [\baddr,\eaddr]
  \]
  Similarly for seals and stack pointers:
  \[
    \range{\seal{\_,\sigma_\baddr,\sigma_\eaddr,\_}} = [\sigma_\baddr,\sigma_\eaddr]
  \]
  and
  \[
    \src{\range{\stkptr{\_,\baddr,\eaddr,\_}} = [\baddr,\eaddr]}
  \]
\end{definition}

@@

The source machine is also a capability machine with memory capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine. I could not recall why we had this, so I removed it. (I think it may have been from when we had call and return instructions?)}
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

\subsection{Operational Semantics}
\subsubsection{Notes}
Genrally:
\begin{itemize}
%\item Opaque capabilities: Only allow locality and permission to be inspected (i.e., no addresses disclosed).
\item Linear capabilities are cleared when they move around in memory.
\end{itemize}

@@

\noindent\textbf{restrict}\\
This instruction uses the $\decPerm{}$ function.
\begin{align*}
  \sem{\trestrict{r_1}{r_2}{r_3}} = &
                                      \begin{cases}
                                        \updPcAddr{}\left(
                                          \arraycolsep=0pt
                                          \begin{array}{rl}
                                          \Phi&\updReg{r_2}{w_2}\\
                                              &\updReg{r_1}{c}
                                          \end{array} \right)
&
                                        \arraycolsep=0pt
                                        \begin{array}[t]{l}
                                          \Phi(r_2) = ((\perm,\lin),\baddr,\eaddr,\aaddr) \tand \\
                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                          \quad\text{and in either case $n \in \ints$ and} \\
                                          \decPerm{n} \sqsubseteq \perm \tand \\
                                          c = ((\decPerm{n},\lin),\baddr,\eaddr,\aaddr) \tand \\Meget
                                          w_2 = \linCons{\Phi(r_2)}
                                        \end{array}
                                        \\
                                        \sourcecolor\left.
                                        \updPcAddr{}\left(
                                          \arraycolsep=0pt
                                          \begin{array}{rl}
                                          \Phi&\updReg{r_2}{0}\\
                                                    &\updReg{r_1}{c}
                                          \end{array}\right)\right.
                                        &
                                        \sourcecolor\left.
                                        \arraycolsep=0pt
                                        \begin{array}[t]{l}
                                          \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                          \text{either $n = r_2$ or $n = \Phi.\reg(r_2)$}\\
                                          \quad\text{and in either case $n \in \ints$ and} \\
                                          \decPerm{n} \sqsubseteq \perm \tand \\
                                          c = \stkptr{\decPerm{n},\baddr,\eaddr,\aaddr}
                                        \end{array}\right.
                                        \\
                                        \failed & \totherwise
                                      \end{cases}
\end{align*}

@@

\noindent\textbf{cseal}\\
\begin{align*}
  \sem{\tcseal{r_1}{r_2}{r_3}}(\Phi) = &
                                  \begin{cases}
                                    \updPcAddr{}\left(
                                    \arraycolsep=0pt
                                    \begin{array}{rl}
                                      \Phi&\updReg{r_1}{\vsc}\\
                                          &\updReg{r_2}{w_2}
                                    \end{array}\right)
&
                                    \arraycolsep=0pt
                                    \begin{array}[t]{l}
                                      \Phi(r_2) \in \SealableCaps \tand \\
                                      \Phi(r_3) = \seal{\lin, \sigma_\baddr, \sigma_\eaddr,\sigma} \tand \\
                                      \sigma_\baddr \leq \sigma \leq \sigma_\eaddr \tand \\ 
                                      \vsc = \sealed{\Phi(r_2),\sigma} \tand\\
                                      w_2 = \linCons{\Phi(r_2)}                                     
                                    \end{array}
                                    \\
                                    \failed & \totherwise
                                  \end{cases}
\end{align*}

@@

We would like splice and split to have following properties

@@

\subsection{Programs and contexts}
\begin{definition}
  \label{def:self-contained}
  \begin{itemize}
  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$
    are self-contained iff
  \[
    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom{\ms}
  \]
  and
  \begin{multline*}
    \forall a \in \dom(\ms) \ldotp\\
    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom{\ms}
  \end{multline*}
  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$ are self contained with slack $\ms_s$ iff
  \[
    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom{\ms \cup \ms_s}
  \]
  and
  \begin{multline*}
    \forall a \in \dom(\ms) \ldotp\\
    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom{\ms \cup \ms_s}
  \end{multline*}
\end{itemize}
\end{definition}

\subsubsection{Programs}
Here we define what consitutes a program.

\begin{definition}
  \label{def:program}
  \lau{27-09-2017: The ``somehow''s need to be fixed.}
  A program consists of two sealed capabilities, $\vsc_\code = \sealed{c_\code,\sigma}$ and $\vsc_\data = \sealed{c_\data,\sigma}$, and and two memory segments $\ms$ and $\ms_\link$ satisfying
  \begin{itemize}
  \item $c_\code = ((\rx,\lin_\code),\baddr_\code,\eaddr_\code,\aaddr_\code)$ where $\aaddr = \baddr + 3$ and $[\baddr_\code,\eaddr_\code]\subseteq \dom(\ms)$
  \item $c_\data = ((\rw,\lin_\data),\baddr_\data,\eaddr_\data,\baddr_\data)$ where $[\baddr_\data,\eaddr_\data] \subseteq \dom(\ms)$ and $[\baddr_\code,\eaddr_\code] \cap [\baddr_\data,\eaddr_\data] = \emptyset$
  \item $\ms$ and $[c_\code,c_\data]$ are self-contained with $\ms_\link$ as slack.
  \item $\ms(\baddr_\code) = \seal(\normal,\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr)$ where these are somehow the sufficient amount of seals.
  \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
  \item $\ms(\baddr_\code+2) = \dots$ where $\dots$ is somehow the stack height.
  \end{itemize}
\end{definition}
\lau{27-09-2017: At the moment, we do not have the proper support to let the code capability be linear (in the sense that you cannot write a sensible program) - maybe we should just require the code capability to be normal.}

\subsubsection{Contexts}

\section{Compiler}
\[
\comp{\cdot} : \src{i}^* \fun \trg{i}^*
\]
